---
title: "Untitled"
author: "William Yip"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
```


```{r}
here("data", "knb-lter-hfr.109.18")
```

```{r}
plant <- read_csv("~/github/ENVS-193DS_homework-05/data/knb-lter-hfr.109.18/hf109-01-sarracenia.csv") %>%
  #make column names cleaner
  clean_names() %>%
  #selecting columns of interest
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)

```
Visualizing the missing data:

```{r}
gg_miss_var(plant)

#missing observations will be excluded
```
subsetting the data by dropping NAs:

```{r subset-drop-NA}
plant_subset <- plant %>%
  drop_na(sla, chlorophyll, num_lvs, num_phylls, amass)
```

Create a correlation plot:

Example writing: To determine the relationships between numerical values in our dataset, we calculated Pearsons r and visually represented correlation using a correlation plot.

```{r}
plant_cor <- plant_subset %>%
  select(feedlevel:num_phylls) %>%
  cor(method = "pearson")

#create correlation plot
corrplot(plant_cor, 
         method = "ellipse",
         addCoef.col = "black"
         )

```

Create a plot of each variable compared against the others

```{r pair-plot}
plant_subset %>%
  select(species:num_phylls) %>%
  ggpairs()
```

Starting regression here:

Example: To determine how species and physiological characteristics predict biomass, we fit multiple linear models. 

```{r}
null <- lm(totmass ~ 1, data = plant_subset) #start with nothing in there
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset) #everything in there
```

We visually assess normality and homeoelasticity of residuals using diagnostic plot for the full model
```{r full-diagnosis}
par(mfrow = c(2,2))
plot(full)

```

We also tested for normality and heteroscedasticity using the Shapiro-Wilk test (null hypothesis: variable of interest (i.e the residuals) are normally distributed). 

We tested for heteroskedaslicity using the Breush-Pagan test (null hypothesis: variable of interst has constant variance)
```{r}
check_normality(full)
check_heteroscedasticity(full)
```

This dataset does not conform to the assumptions of linear regression (normal for bio datasets)
Use a log10 of each observation to transform the resposne variable to transform residuals to normal

```{r}
full_log <- lm(log(totmass) ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)

plot(full_log)
check_normality(full_log)
check_heteroskedasticity(full_log)
```